Data Exploration: IUCN and AquaMaps

### 2016-05-17 data_explore/data_explore.Rmd

>>>>

``` {r pct_overlap_vs_ratio_of_area2, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

spp_list_quads <- read_csv(file.path(dir_data, 'spp_list_quads.csv'))

ggtheme_plot2 <- ggtheme_basic + 
  theme(panel.border     = element_blank(),
        panel.grid.minor = element_blank(), 
#        panel.grid.major = element_line(colour = 'grey90'),
        panel.grid.major = element_blank(),
        axis.title.x = element_text(hjust = 0),
        axis.title.y = element_text(hjust = 1),
        axis.line = element_blank())

### mongo plot time
scatter_quadplot <- ggplot(spp_list_quads %>% 
                             transform(popn_category = reorder(popn_category, category_score)) %>%
                             mutate(quad = factor(quad, levels = c('q1', 'q2', 'q3', 'q4'))),
                           aes(x = area_ratio, 
                               y = sm_perc, 
                               color = quad)) +
  ggtheme_plot2

### add mean lines (quadrant dividers) under the points
scatter_quadplot <- scatter_quadplot +
  geom_vline(xintercept = area_align_mean,
             color      = 'grey80') +
  geom_hline(yintercept = dist_align_mean,
             color      = 'grey80')

### add the points!
scatter_quadplot <- scatter_quadplot +
  geom_point(alpha = .8) +
  scale_colour_manual(values = c('q1' = '#4dac26',
                                 'q2' = '#b8e186',
                                 'q3' = '#f1b6da',
                                 'q4' = '#d01c8b')) +
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 103),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%'))

### add axis lines and mean labels:
scatter_quadplot <- scatter_quadplot  +
  geom_vline(xintercept = 0,
             color      = 'grey80') +
  geom_hline(yintercept = 0,
             color      = 'grey80') +
  annotate(geom = 'text',
           x = (area_align_mean - .5), y = 5,
           hjust = 0, vjust = 0,
           color = 'grey30', size = 2,
           fontface = 'bold.italic', angle = 90,
           label = sprintf('Mean = %s%%', round(area_align_mean, 1))) +
  annotate(geom = 'text',
           x = 5, y = (dist_align_mean + .5),
           hjust = 0, vjust = 0,
           color = 'grey30', size = 2,
           fontface = 'bold.italic', angle = 0,
           label = sprintf('Mean = %s%%', round(dist_align_mean, 1)))

### quadrant labels
scatter_quadplot <- scatter_quadplot +
  annotate(geom = 'text', color = 'grey30', size = 3, fontface = 'bold.italic',
           x = 2, y = 2,
           hjust = 0, vjust = 0,
           label = 'Poorly aligned') +
  annotate(geom = 'text', color = 'grey30', size = 3, fontface = 'bold.italic',
           x = 98, y = 98,
           hjust = 1, vjust = 1,
           label = 'Well aligned') +
  annotate(geom = 'text', color = 'grey30', size = 3, fontface = 'bold.italic',
           x = 2, y = 98,
           hjust = 0, vjust = 1,
           label = 'Distribution aligned') +
  annotate(geom = 'text', color = 'grey30', size = 3, fontface = 'bold.italic',
           x = 98, y = 2,
           hjust = 1, vjust = 0,
           label = 'Area aligned')

scatter_quadplot <- scatter_quadplot +
  labs(x = bquote('Area ratio '*alpha['ext']*''), 
       y = bquote('Distribution alignment '*alpha['dist']*'')) +
  theme(legend.position = 'none')

ggsave(file.path(dir_fig, 'scatter_quadmap_colbyquad.png'),
       width = 17.8, height = 9, units = 'cm')
